Tree size, (microhabitat)?, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-09-09 manuscript/climate_sensitivity_draft.Rmd
added gypsy moth info

*Descriptions of Droughts*
In the 60-year period between 1950 and 2009, there were three droughts that met our criteria of anomalously dry climatic conditions coupled with substantial reductions in tree growth for at least some portions of the community: 1966, 1977, and 1999 (Fig. 1). 
We excluded one year (1991) meeting the growth reduction criteria (26.5% of trees experienced >30% growth reduction, mean resistance= -13.8%) because this year was not among the strongest droughts of the study period (**DETAILS**). Rather, reduced growth may be explained in part because gypsy moths (*Lymantria dispar L.*) were present in the region from XXXX-XXXX, causing extensive defoliation of *Quercus* spp. (**#% of cores/propoductivity**), with peak impact at SCBI within 3 years prior to 1991 (**Cite Shenanadoah paper, if accepted**). 
Climatically, these droughts included three of the five years between 1950 and 2009 with greatest moisture deficit (PET-PRE) during the peak growing season months of May-August, which are the months to which annual tree growth at this site is most sensitive [@helcoski_growing_2019]. Specifically, 1966, 1977, and 1999 had mean MJJA PET-PRE of 83.37, 86.97, and 80 mm mo<sup>-1</sup>, respectively. The years 1964 and 2007 also ranked in the top five driest (PET-PRE =83.87 and 82.13 mm mo<sup>-1</sup>), but *were not among the lowest in terms of PDSI and were not identified as a pointer yeasr.* 
**The droughts differed in timing/duration/etc. .. The year 1966 was preceded by two relatively dry years... 1964 among five driest in terms of May-August [PET-PRE],  1965  also anomalously hot and dry. **

>>>>

*Descriptions of Droughts*
In the 60-year period between 1950 and 2009, there were three droughts that met our criteria of anomalously dry climatic conditions coupled with substantial reductions in tree growth for at least some portions of the community: 1966, 1977, and 1999 (Fig. 1). 
We excluded one year (1991) meeting the growth reduction criteria (26.5% of trees experienced >30% growth reduction, mean resistance= -13.8%) because this year was not among the strongest droughts of the study period (**DETAILS**). Rather, the severity of growth reduction may be explained by defoliation by gypsy moths (*Lymantria dispar L.*) from approximately 1988-1995, which most stronly impacted *Quercus* spp. (**#% of cores/propoductivity**) (**Cite Shenanadoah paper, if accepted**). 
Climatically, these droughts included three of the five years between 1950 and 2009 with greatest moisture deficit (PET-PRE) during the peak growing season months of May-August, which are the months to which annual tree growth at this site is most sensitive [@helcoski_growing_2019]. Specifically, 1966, 1977, and 1999 had mean MJJA PET-PRE of 83.37, 86.97, and 80 mm mo<sup>-1</sup>, respectively. The years 1964 and 2007 also ranked in the top five driest (PET-PRE =83.87 and 82.13 mm mo<sup>-1</sup>), but *were not among the lowest in terms of PDSI and were not identified as a pointer yeasr.* 
**The droughts differed in timing/duration/etc. .. The year 1966 was preceded by two relatively dry years... 1964 among five driest in terms of May-August [PET-PRE],  1965  also anomalously hot and dry. **
